---
title: "eth_btc_retroactive"
author: "Charliemarketplace"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_float:
      collapsed: false
editor_options:
  chunk_output_type: console
---

# Intro

This is a Profit and Loss calculation including potential divergent loss (DL) of all closed positions in the ETH-WBTC 0.3% Uniswap v3 pool from inception of the pool up to the  block height of 15,576,600 (September 20, 2022). It reviews *all* changes in liquidity and accumulated fees to identify the profit and loss of every unique position across 3 accountings: 

- Direct Profit & Loss, accounting for price changes in the underlying assets
-  Opportunity Cost, i.e.,  HODL Reference Value: what assets deposited would have been worth as of the block where the position was closed.
- Realized Value, i.e., Strategy Reference Value: what the assets withdrawn were worth as of the block where the position was closed including accumulated fees.

Direct P&L is useful for understanding gains in the context of changing prices, here, P&L is measured in both ETH and BTC terms. 

Divergent Loss (sometimes called Impermanent Loss) is the difference between the results of using a strategy (here, fees and automatic market making in Uniswap v3) and the counter factual: the amount of assets if one were *not* to have used the strategy. 
A more in depth view on definitions and calculations is available in the ETH USDC Retroactive over the same time range.

# Data Collection 

The `collect_ethbtc_data.R` script included in this repo uses the Flipside Crypto's shroomDK API to pull all the relevant data. For brevity, this markdown reads from a saved RDS copy *not* available in the repo. To reproduce this analysis, you can run the `collect_ethbtc_data.R` script using your own shroomDK API key available for free.

```{r}
library(gmp) # large numbers
library(reactable) # clean tables
library(dplyr) # data manipulation
options(scipen = 10)

source("../key_functions.R")

lp_actions <- readRDS("ethbtc_lp_actions.rds")

# remove actions that don't remove liquidity, these are fee collection events
lp_actions <- lp_actions %>% filter(LIQUIDITY != 0)

fees <- readRDS("ethbtc_fees.rds")
btc_prices <- readRDS("ethbtc_prices.rds")
btc_prices <- btc_prices[,c("BLOCK_NUMBER", "BTC_MARKET_PRICE")]
colnames(btc_prices) <- c("BLOCK_NUMBER","btc_price")

```

### Closed Positions at Block Height

```{r}
lp_actions$unique_id <- lp_actions$NF_TOKEN_ID

custom_index <- which(is.na(lp_actions$unique_id))

lp_actions[custom_index, "unique_id"] <- paste0(lp_actions[custom_index, "NF_POSITION_MANAGER_ADDRESS"],
                                                "--",
                                                lp_actions[custom_index, "TICK_LOWER"],
                                                "--",
                                                lp_actions[custom_index, "TICK_UPPER"])

# adding BTC Market Price from btc_prices to lp_actions via merge for later 
lp_actions <- merge(lp_actions, y = btc_prices, all.x = TRUE, all.y = FALSE, by = "BLOCK_NUMBER")

# manual insertion of first price 
lp_actions$btc_price[is.na(lp_actions$btc_price)] <- 16.07226

# while here do fees too
fees$unique_id <- fees$NF_TOKEN_ID
custom_index <- which(is.na(fees$unique_id))
fees[custom_index,"unique_id"] <- paste0(fees[custom_index, "NF_POSITION_MANAGER_ADDRESS"],
                                                "--",
                                                fees[custom_index, "TICK_LOWER"],
                                                "--",
                                                fees[custom_index, "TICK_UPPER"])

# also while here, ensure only 1 fee collection row per tx_hash, i.e., just add them

fees <- fees %>% 
  group_by(across(c(-AMOUNT0_ADJUSTED, -AMOUNT1_ADJUSTED))) %>% 
  summarise(
  AMOUNT0_ADJUSTED = sum(AMOUNT0_ADJUSTED),
  AMOUNT1_ADJUSTED = sum(AMOUNT1_ADJUSTED), .groups = "drop") %>%  
  relocate(c(NF_TOKEN_ID, AMOUNT0_ADJUSTED, AMOUNT1_ADJUSTED), .after = NF_TOKEN_ID) %>% 
  as.data.frame()


liquidity_at_timestamp <- lp_actions %>% mutate(
  liquidity_signed = ifelse(ACTION == "DECREASE_LIQUIDITY", LIQUIDITY * -1, LIQUIDITY)
) %>% group_by(unique_id) %>% 
  summarise(sumliq = sum(liquidity_signed))

# due to some large number precision errors, some liquidity may be technically negative 
# but effectively 0 for our purposes.
closed_positions <- liquidity_at_timestamp %>% filter(sumliq <= 0)

```

There are `r nrow(closed_positions)` ETH-BTC 0.3% closed positions, including vaults, as of Sept 20, 2022.

## Functions

```{r}
price_col = "btc_price"

pnl <- function(id_accounting, t0col, t1col, price_col, price_base = "t1/t0"){
  
  # allow user to choose the base that is easier to read
  # e.g., usdc/eth may be easier to read than eth/usdc; which means t0/t1 instead of t1/t0
  # while ETH/BTC may be easier than BTC/ETH; which is t1/t0 the default!
  
  if(price_base == 't1/t0'){
     pnl = data.frame(
      pnl_t0_terms = sum(id_accounting$token0, (id_accounting$token1/id_accounting[[price_col]])),
      pnl_t1_terms = sum(id_accounting$token1, (id_accounting$token0*id_accounting[[price_col]]))
    )
  } else if(price_base == 't0/t1'){
    pnl = data.frame(
      pnl_t0_terms = sum(id_accounting$token0, (id_accounting$token1*id_accounting[[price_col]])),
      pnl_t1_terms = sum(id_accounting$token1, (id_accounting$token0/id_accounting[[price_col]]))
    )
  } else {
    stop("price_base must be 't1/t0' (default) or 't0/t1'")
  }
 

colnames(pnl) <- c(t0col, t1col)

return(pnl)
}

hodl_reference <- function(id_accounting, t0col, t1col, price_col, price_base = "t1/t0"){
  
  price_at_close_block = id_accounting[[price_col]][id_accounting$BLOCK_NUMBER == max(id_accounting$BLOCK_NUMBER)][1] 
  
  if(price_base == 't1/t0'){
    hodl <- id_accounting %>% 
      filter(accounting == 'cost_basis') %>% 
      summarise(
        hodl_t0_terms = abs(sum(token0) + sum(token1 / price_at_close_block)),
        hodl_t1_terms = abs(sum(token1) + sum(token0 * price_at_close_block)),
      )
  } else if(price_base == 't0/t1'){
    hodl <- id_accounting %>% 
      filter(accounting == 'cost_basis') %>% 
      summarise(
        hodl_t0_terms = abs(sum(token0) + sum(token1 * price_at_close_block)),
        hodl_t1_terms = abs(sum(token1) + sum(token0 / price_at_close_block)),
      )
  } else {
    stop("price_base must be 't1/t0' (default) or 't0/t1'")
  }
  
  colnames(hodl) <- c(t0col, t1col)
  
  return(hodl)
}

strategy_reference <- function(id_accounting, t0col, t1col, price_col, price_base = "t1/t0"){
  
  price_at_close_block = id_accounting[[price_col]][id_accounting$BLOCK_NUMBER == max(id_accounting$BLOCK_NUMBER)][1] 
  
  if(price_base == 't1/t0'){
    strat <- id_accounting %>% 
      filter(accounting != 'cost_basis') %>% 
      summarise(
        strategy_t0_terms = abs(sum(token0) + sum(token1 / price_at_close_block)),
        strategy_t1_terms = abs(sum(token1) + sum(token0 * price_at_close_block)),
      )
  } else if(price_base == 't0/t1'){
    strat <- id_accounting %>% 
      filter(accounting != 'cost_basis') %>% 
      summarise(
        strategy_t0_terms = abs(sum(token0) + sum(token1 * price_at_close_block)),
        strategy_t1_terms = abs(sum(token1) + sum(token0 / price_at_close_block)),
      )
  } else {
    stop("price_base must be 't1/t0' (default) or 't0/t1'")
  }
  
  colnames(strat) <- c(t0col, t1col)
  return(strat)
  
}


```

# Fee Correction

```{r}

closed_fees <- fees %>% 
  filter(unique_id %in% closed_positions$unique_id) %>% 
  arrange(BLOCK_NUMBER)

closed_lp_actions <- lp_actions %>% 
  filter(unique_id %in% closed_positions$unique_id)

closed_lp_w_fees <- closed_lp_actions %>% filter(
  unique_id %in% closed_fees$unique_id
) %>% arrange(BLOCK_NUMBER) %>% 
  select(BLOCK_NUMBER, TX_HASH, unique_id, ACTION, AMOUNT0_ADJUSTED, AMOUNT1_ADJUSTED, btc_price)


# split automatically alphabetizes by ID # 
id_fees_list <- split(closed_fees, f = closed_fees$unique_id)
id_lp_list <- split(closed_lp_w_fees, f = closed_lp_w_fees$unique_id)

# If the order of IDs don't exactly match note the problem 
if(
   ! identical(names(id_lp_list), names(id_fees_list))
) {
  stop("List of IDs are not aligned")
}

# otherwise run through the list of fees and correct them 
# NOTE fee_correction only safe to use when split by unique_id
# Possible for a tx_hash to include multiple position interactions

id_fees_fixed <- lapply(1:length(id_fees_list), 
                        FUN = function(i){
                          fee_correction(id_fees_list[[i]], id_lp_list[[i]])
                        })
```

### Accounting 

```{r}
aa <- accounting(id_lp_actions = id_lp_list[[100]], 
           id_fees = id_fees_fixed[[100]], 
           price_col = "btc_price")

pnl(aa, "pnl_btc_terms", "pnl_eth_terms","btc_price")
hodl_reference(aa, "hodl_btc_terms", "hodl_eth_terms","btc_price")
strategy_reference(aa, "strat_btc_terms", "strat_eth_terms","btc_price")

```

```{r}

id_accounting <- lapply(1:length(id_lp_list), 
                        FUN = function(i){
                          tryCatch(
                          accounting(id_lp_actions = id_lp_list[[i]],
                                     id_fees = id_fees_fixed[[i]],price_col = "btc_price"),
                          error = function(e){
                              print(paste0("Error found in position: ", i))
                            })  
                        })

names(id_accounting) <- names(id_lp_list)

# to simplify segmentation of analysis saving this
saveRDS(id_accounting, "post_ethbtc_accounting.rds")

# for brevity read id_accounting to continue here
id_accounting <- readRDS("post_ethbtc_accounting.rds")

id_pnl <- lapply(id_accounting, function(i){
  pnl(i, "pnl_btc_terms", "pnl_eth_terms","btc_price")
  })
pnl_results <- bind_rows(id_pnl, .id = "unique_id")

id_hodl <- lapply(id_accounting, function(i){
  hodl_reference(i, "hodl_btc_terms", "hodl_eth_terms","btc_price")
})

hodl_results <- bind_rows(id_hodl, .id = "unique_id")

id_strat <- lapply(id_accounting, function(i){
  strategy_reference(i, "strat_btc_terms", "strat_eth_terms","btc_price")
})

strategy_results <- bind_rows(id_strat, .id = "unique_id")

ethbtc_results <- merge(pnl_results, hodl_results, by = "unique_id")
ethbtc_results <- merge(ethbtc_results, strategy_results, by = "unique_id")

saveRDS(ethbtc_results, file = "ethbtc_results.rds")


```
